---
title: "bilingual R Markdown"
author: "theophano mitsa"
date: "August 1, 2019"
output: html_document
---

```{r setup, include=FALSE}
  library(reticulate)
use_python("C:/Users/Theo/Anaconda3/python.exe")
knitr::opts_chunk$set(echo = TRUE)
```

Discover the Python that will be used
```{r}
py_discover_config()
```
1. DATAFRAME CREATION, SUMMARIZATION

a. Read a dataframe from scratch
IN R

```{r}
library(tidyverse)
set.seed(3)
mat1<-matrix(rnorm(6,12,3), nrow=2)
dimnames(mat1)=list(c("dep1","dep2"),c("Tom","Jim","Sam"))
DepositFrame=as.data.frame(mat1)
print(DepositFrame)

```

IN PYTHON

```{python}
import numpy as np
import pandas as pd
ChildAgeFrame = pd.DataFrame(np.array(([15, 0],[69,3], [35, 2])),
               columns=['age', 'children'],                                 index=['Tes', 'Linda', 'Kate'])
print(ChildAgeFrame)
ChildAgeFrame.describe()
#Print the Tes column
print(ChildAgeFrame.loc['Tes'])
#Alternate way to create a DataFrame
HotelFrame = pd.DataFrame({'Room Type': ['Regular','Suite', 'Regular','Lux Suite','Suite'],'Floor':['1','2','3','3','2']})
print(HotelFrame)
```
b. Read a dataframe from a csv file

IN R
```{r}
heart_R<-read.csv("heart.csv",header=TRUE, fileEncoding="UTF-8-BOM")
#print dimensions
dim(heart_R)
#print names of columns
names(heart_R)
```

IN PYTHON
```{python}
import pandas as pd
heart_P=pd.read_csv('heart.csv')
#read only certain columns from a csv file
drop_cols=['thal','thalach']
heart_Pr=pd.read_csv('heart.csv',usecols=lambda cc : cc not in drop_cols, index_col=False)
#print dimensions
print(heart_P.shape)
#print names of columns
print(heart_P.columns)
```

c. Data Summarization

In R
```{r}
#print first rows
head(heart_R)
#Print some basic info about the dataframe
glimpse(heart_R)
str(heart_R)
summary(heart_R)
#Alternate way to summarize variables
library(prettyR)
describe(heart_R)
#sort according to cholesterol
heart_R%>% arrange(chol)%>%head()
#Use tapply() to group and compute functions within groups
tapply(heart_R$chol,heart_R$sex,mean)
#compute the mean after grouping with group_by()
heart_R%>%group_by(sex)%>%summarise_all(~mean(.,na.rm=TRUE))
#Compute counts of groups
heart_R%>%count(sex)
heart_R%>%count(cp,sort=TRUE)
heart_R%>%count(cp,fbs, sort=TRUE)

```
In Python

```{python}
import pandas as pd
heart_P=pd.read_csv('heart.csv')
#print first rows
print(heart_P.head(5))
#Print some basic info about the dataframe
heart_P.describe()
#Count grouped cases
heart_P.groupby('fbs').count()
#Count the mean of grouped cases
heart_P.groupby(['sex','fbs']).count()
heart_P.groupby(['sex','fbs']).mean()
```


2. INDEX-BASED SELECTION: SELECT ROWS,COLUMNS, ELEMENTS USING INDICES

IN R
```{r}
#Print first row
print(DepositFrame[1,])
#Print first column
print(DepositFrame[,1])
#Print first element
DepositFrame[1,1]
```
IN PYTHON

```{python}
import numpy as np
import pandas as pd
ChildAgeFrame = pd.DataFrame(np.array(([15, 0], [69, 3],[35, 2])),
               columns=['age', 'children'],                                 index=['Tes', 'Linda', 'Kate'])
#Select first row. Returns Series.
print(ChildAgeFrame.iloc[0])
#Select first row. Returns Dataframe.
print(ChildAgeFrame.iloc[[0]])
#Select first column
print(ChildAgeFrame.iloc[:,0])
#Select last column
print(ChildAgeFrame.iloc[:,-1])
#Select first element
print(ChildAgeFrame.iloc[0,0])
#Select first 2 rows
print(ChildAgeFrame.iloc[0:2])
#Select first 2 columns
print(ChildAgeFrame.iloc[:,0:2])

```
3.NON-INDEX-BASED SELECTION

a. Column Selection Using Column Names

IN R
```{r}
#selecting a single column in basic R
DepositFrame["Tom"]
#selecting a multitude of columns using select()
DepositFrame%>%select("Tom","Sam")
```

IN PYTHON
```{python}
import numpy as np
import pandas as pd
ChildAgeFrame = pd.DataFrame(np.array(([15, 0], [69, 3], [35,2])),
               columns=['age', 'children'],                                 index=['Tes', 'Linda', 'Kate'])

print(ChildAgeFrame[['age']])
```

b.Column Selection Using Regular Expressions, Logical Conditions

IN R
```{r}
#Filter using regex
 DepositFrame%>%select(matches("S.+m"))%>%glimpse()
#Filter based on a logical expression
DepositFrame %>% select(matches("m"))%>%
  select_if(~max(.,na.rm=TRUE)<12)%>%glimpse()
```

IN PYTHON
```{python}
import numpy as np
import pandas as pd
ChildAgeFrame = pd.DataFrame(np.array(([15, 0], [69, 3], [39,2])),
               columns=['age', 'children'],                                 index=['Tes', 'Linda', 'Kate'])
Child_select=ChildAgeFrame.filter(regex='a.+e',axis=1)
print(Child_select)
```

3.c Non-Index-Based Row Selection: With Logical Expressions EXPRESSIONS And/Or Row Names

IN R
```{r}
#Print the age of men with high cholesterol
#With basic R
MEN_WITH_HIGH_CHOL<-heart_R$chol > 280 & heart_R$sex==1
print(heart_R[MEN_WITH_HIGH_CHOL,1])
#with dplyr
filtered<-filter(heart_R,chol > 280 & sex==1)
print(filtered[1])
#print any rows that contain a value greater than 280
heart_n<-heart_R%>%filter_all(any_vars(.>280))
print(heart_n)
```

IN PYTHON
```{python}
import numpy as np
import pandas as pd
#Select row using logical condition
heart_P=pd.read_csv('heart.csv')
df=heart_P[(heart_P['chol']> 200) & (heart_P['thalach']==178)]
print(df)

#Select row using row name
dfnew = pd.DataFrame(np.array(([15,0], [69,3],[35,2])),
               columns=['age', 'children'],                                 index=['Tes', 'Linda', 'Kate'])
print(dfnew.loc['Linda'])
```

4. DELETE COLUMNS/ROWS


IN R
```{r}
#Delete using column name
print(DepositFrame)
Depnew<-DepositFrame%>%select(-c("Sam"))
Depnew
#Delete using starts_with.
Depnew<-DepositFrame%>%select(-starts_with("J"))
Depnew
#Delete row
set.seed(3)
mat1<-matrix(rnorm(6,12,3), nrow=2)
dimnames(mat1)=list(c("dep1","dep2"),c("Tom","Jim","Sam"))
mat1
df=as.data.frame(mat1)
names<-rownames(df)
names
todrop=names[1]
todrop
df2<-df[!(rownames(df)%in%todrop),]
print(df2)


```


IN PYTHON

```{python}
import pandas as pd
HotelFrame = pd.DataFrame({'Room Type': ['Regular','Suite', 'Regular','Lux Suite','Suite'],'Floor':['1','2','3','3','2']})
#Delete a column using its name
Hotelnew=HotelFrame.drop('Floor', axis=1)
print(Hotelnew)

#Alternate way
Hotelnew=HotelFrame.drop(columns=['Floor'],axis=1)
print(Hotelnew)

#Delete Row
Hotelnew=HotelFrame.drop([0])
print(Hotelnew)
```
5. ADD ROW/ COLUMN 

IN R
```{r}
#ADD COLUMN
DepositFrame$Brandon<-c(9,12)
#Using mutate, transmute. Transmute only keeps the new vars
dfnew<-mutate(DepositFrame,sumdep=Tom+Jim, Sam=NULL)
print(dfnew)
#ADD ROW
addit_row<-data.frame('Tom'=11.0, 'Jim'=12.2,'Sam'=10.2,'Brandon'=11.3)
df2<-rbind(DepositFrame,addit_row)
df2
#Combine group by and mutate to do calculations within groups.

heart_Rn<-heart_R%>% group_by(sex)%>%mutate(chol_n=chol/mean(chol,na.rm=TRUE))
print(heart_Rn)

```


IN PYTHON
```{python}
import numpy as np
import pandas as pd
HotelFrame = pd.DataFrame({'Room Type': ['Regular','Suite', 'Regular','Lux Suite','Suite'],'Floor':['1','2','3','3','2']})
pets=['No','Yes','No','No','Yes']
HotelFrame['pet']=pets
#Using a map to implement logic
pets_allowed={'Regular':'No','Suite':'Yes','Lux Suite':'No'}
#Add column using the above map
HotelFrame['PETS']=HotelFrame['Room Type'].map(pets_allowed)
#ADD a ROW
dfadd=HotelFrame.append({'Room Type':'Regular','Floor':'1',
'pet':'No','PETS':'No'}, ignore_index=True)
print(dfadd)
```

6. APPLY A FUNCTION TO ROWS/COLUMNS OF A DATAFRAME

IN R
```{r}
#Compute mean over columns
apply(DepositFrame,2,mean)
#Compute mean over rows
apply(DepositFrame,1,mean)
```
IN PYTHON
```{python}
import numpy as np
import pandas as pd
ChildAgeFrame = pd.DataFrame(np.array(([15, 0], [69,3],[35,2])),
               columns=['age', 'children'],                                 index=['Tes', 'Linda', 'Kate'])
#Compute the mean of the columns
print(ChildAgeFrame.apply(np.mean,axis=0))
#Apply a lambda function to all elements
print(ChildAgeFrame.apply(lambda x: x*5))

```

7. SAMPLING A DATAFRAME

IN R
```{r}
set.seed(5)
#Sample 3 rows without replacement
heart_R %>% sample_n(3, replace = FALSE)

# Sample 1% of rows, without replacement
heart_R %>% sample_frac(0.01, replace = FALSE)
```

IN PYTHON
```{python}
import numpy as np
import pandas as pd
heart_P=pd.read_csv('heart.csv')
s1=heart_P.sample(n=3)
s2=heart_P.sample(frac=0.01)
```

8. COMPARE DATA LOADING BETWEEN DATA.TABLE AND DATAFRAME

```{r}
library(data.table)
system.time(student1<-read.csv("student_data.csv",header=TRUE, fileEncoding="UTF-8-BOM"))
system.time(student2<-fread("student_data.csv"))

```

9. COMPARE FILTERING IN DPLYR AND DATA.TABLE
```{r}
library(data.table)
heart_RDT=fread("heart.csv",header=TRUE)
ff<-heart_R%>%filter(chol > 280 & sex==1)
#Mean age of men with chol > 280
mean(ff[,1])
#Mean age computed from a data.table
dd<-heart_RDT[chol > 280 & sex==1,.(mm=mean(age))]
dd
```

10. IMPLEMENT/COMPARE PARALLELIZATION/ FUNCTION COMPILATION
```{r}
library(parallel)
library(compiler)
k<-detectCores()
c1<-makeCluster(4)
mysquare<-function(num){return(num^2)}
#Compile the function
gb<-cmpfun(mysquare)
set.seed(2)
sample_numbers <- sample(10000000, 10000000)
system.time(sapply(sample_numbers,mysquare))
system.time(sapply(sample_numbers,gb))
system.time(parSapply(c1,sample_numbers, mysquare))
system.time(parSapply(c1,sample_numbers, gb))
stopCluster(c1)

```


